---
title: "Civic Database Cancer Genes"
author: "jl"
date: "19/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Outline

This code extracts genes with their mutations and cancer types from a Civic JSON format database, and saves the results to a CSV file.

### R Packages Required

There are a few packages that need to be installed to run the R code within this document. These are primarily in or adjacent to the `tidyverse` set of [packages](https://www.tidyverse.org/packages/) and consist of the following:

* The `tidyverse` set of packages;
* The `jsonlite` package for `JSON` format data;
* The `knitr` package for report generation.

Each of these can be installed with `install.packages("<name_of_package>")` at the `R` console, or via RStudio's `Tools` menu with `Install Packages`.

In all of the following `R` code snippets, it is implied that the `tidyverse` library has been loaded. For clarity this is done here, rather than invisibly in the setup chunk, with:

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

Other libraries that are used in only a few code chunks, are explicitly loaded in those chunks for clarity.

# Process Civic Database

## Load Civic Database from local JSON file

```{r, eval=FALSE}
library(jsonlite)

fname <- 'db/civic.json'
cat("reading from JSON file: ", fname, "\n")
rawj <- fromJSON(fname)

# extract just the variants[] as a dataframe
variants <- rawj$variants
```

## Extract nested data into dataframe

```{r}
cancer_gene_types <- variants %>%
  # we want the gene, mutation and the evidence_items
  select(entrez_name, name, evidence_items) %>%
  # note that name NEEDS to be renamed prior to unnest
  # due to the presence of 'name' field within the nested evidence_items
  rename(gene=entrez_name, mutation=name) %>%
  # we use evidence_items: disease, evidence_level, evidence_direction
  unnest_wider(evidence_items) %>%
  # only want disease: report_type
  hoist(disease, cancer_type="report_type") %>%
  # just keep the fields we want for the CSV
  select(gene, mutation, cancer_type, evidence_level, evidence_direction) %>%
  # these columns have the same number of elements - unnest into rows
  unnest(c(cancer_type, evidence_level, evidence_direction))
```

## output dataframe to csv file

```{r}
write_csv(cancer_gene_types, "genes_cancers_list.csv")
```
